Decentralisation of Poverty - Replication Analysis

### 2016-08-05 markdown/draft.RmD
update script, tidying up and automating bayesian model run functions

Quantile   | RCI, 2004 | RCI, 2012  | Change  
---------- | --------  | ---------- | -------  
median     | 0.144     | 0.093      | -0.051
2.5% CrI   | 0.142     | 0.091      | -0.054
97.5% CrI  | 0.146     | 0.095      | -0.048

@@

@@

The following plot shows scatterplots of the ID proportions in each Scottish datazone in 2004, 2009 and 2012. Within these plots the blue line indicates parity between values on the x and y axis, and the red dashed line shows a nonlinear smoother function between the two ID proportions being plotted, and hence the level of deviation between the actual relationship over the two years and parity. 
```{r show_temporal_dep, echo = FALSE, warning = FALSE, EVAL = TRUE}

@@

The following plot shows the change in the total population size as a function of distance from city centre. (The smoother line at distances less than 1km should likely be ignored for most cities as it involves extrapolation based on very few observations). 
```{r plot_changetotalpop, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
totalpopchange_plot_for_city <- function(this_city){
  simd_2011_reweighted  %>% 
    ungroup() %>%  
    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
    filter(nearest_centre == this_city) %>%
    distinct() %>% 
    select(dz_2011, distance_to_centre, year, pop_total) %>% 
    mutate(distance_to_centre = distance_to_centre / 1000) %>% 
    spread(year, pop_total) %>% 
    mutate(change = (`2012` / `2004`) - 1) %>% 
    ggplot(., aes(x = distance_to_centre, y = change)) + 
    geom_point( alpha = 0.01) + 
    scale_x_log10(limits = c(0.5, 50), breaks = c(0.5, 1, 2, 5, 10, 20, 50)) + 
    scale_y_continuous(limits = c(-.5, .5)) +    
    stat_smooth() + 
    geom_vline(aes(xintercept = 1e+0)) + 
    geom_vline(aes(xintercept = 1.2e+1)) +
    geom_hline(aes(yintercept = 0), linetype = "dashed") + 
    labs(x = "Distance to centre (km)", y = "Change in total pop",
         title = this_city)
}

plot_grid(
  totalpopchange_plot_for_city("Aberdeen"),
  totalpopchange_plot_for_city("Dundee"),
  totalpopchange_plot_for_city("Edinburgh"),
  totalpopchange_plot_for_city("Glasgow"),
  nrow = 2
)
```
This plot suggests that there have been substantial increases in population sizes in Aberdeen, Edimburgh and Glasgow over this period, in datazones closer to the city centre, with the greatest changes occurring near the centre of Glasgow. In Aberdeen, and to a lesser extent for Dundee, there have also been increases in population shares living at distances of 12km and more from the city, suggesting greatest consistency in population sizes at intermediate distances.

@@

Alternative methods of reweighting SIMD values released at 2001 datazone levels could be explored, including those which make use of more sophisticated GIS-based methods to apportion values to 2011 datazones, and of approaches which use SIMD component values available at a higher level of spatial disaggregation (e.g. output areas) to recalculate and reapportion SIMD scores to 2011 datazone geographies. As with all analyses, the additional benefit of alternative and more sophisticated approaches should be weighted against the additional time cost involved, and a judgement made about whether the ratio of marginal benefits to marginal costs represents a cost effective use of researcher time.

>>>>

Quantile   | RCI, 2004 | RCI, 2012  | Change  
---------- | --------  | ---------- | -------  
median     | 0.144     | 0.093      | -0.051
2.5% CrI   | 0.142     | 0.091      | -0.054
97.5% CrI  | 0.146     | 0.096      | -0.047

@@

#### Dundee

Quantile   | RCI, 2004 | RCI, 2012  | Change  
---------- | --------  | ---------- | -------  
median     | 0.304     | 0.256      | -0.048
2.5% CrI   | 0.298     | 0.251      | -0.056
97.5% CrI  | 0.309     | 0.262      | -0.040

Quantile   | D, 2004 | D, 2012  | Change  
---------- | ------- | -------- | ------- 
median     | 0.414   | 0.377    | -0.037
2.5% CrI   | 0.409   | 0.372    | -0.044
97.5% CrI  | 0.418   | 0.382    | -0.030

#### Aberdeen

Quantile   | RCI, 2004 | RCI, 2012  | Change  
---------- | --------  | ---------- | -------  
median     | 0.218     | 0.187      | -0.031
2.5% CrI   | 0.213     | 0.181      | -0.040
97.5% CrI  | 0.224     | 0.193      | -0.022

Quantile   | D, 2004 | D, 2012  | Change  
---------- | ------- | -------- | ------- 
median     | 0.385   | 0.360    | -0.024
2.5% CrI   | 0.379   | 0.355    | -0.032
97.5% CrI  | 0.390   | 0.366    | -0.017

#### Edinburgh

Quantile   | RCI, 2004 | RCI, 2012  | Change  
---------- | --------  | ---------- | -------  
median     | 0.013     | -0.013     | -0.025
2.5% CrI   | 0.009     | -0.016     | -0.030
97.5% CrI  | 0.016     | -0.009     | -0.020

Quantile   | D, 2004 | D, 2012  | Change  
---------- | ------- | -------- | ------- 
median     | 0.385   | 0.319    | -0.067
2.5% CrI   | 0.382   | 0.316    | -0.071
97.5% CrI  | 0.389   | 0.322    | -0.062

<!-- > round(apply(indicators_edinburgh, 2, quantile, c(0.5, 0.025, 0.975)),3) -->
<!--       RCI2004 RCI2012 D2004 D2012 -->
<!-- 50%     0.013  -0.013 0.385 0.319 -->
<!-- 2.5%    0.009  -0.016 0.382 0.316 -->
<!-- 97.5%   0.016  -0.009 0.389 0.322 -->
<!-- > ## Differences in RCI and D in 2011 - 2001 -->
<!-- > round(quantile(indicators_edinburgh[ ,2] - indicators_edinburgh[ ,1], c(0.5, 0.025, 0.975)),3) -->
<!--    50%   2.5%  97.5%  -->
<!-- -0.025 -0.030 -0.020  -->
<!-- > round(quantile(indicators_edinburgh[ ,4] - indicators_edinburgh[ ,3], c(0.5, 0.025, 0.975)),3) -->
<!--    50%   2.5%  97.5%  -->
<!-- -0.067 -0.071 -0.062  -->
<!-- >  -->
<!-- > indicators_aberdeen <- do_model("Aberdeen") -->
<!-- Keys match perfectly. -->

<!-- Keys match perfectly. -->

<!-- Joining by: "dz_2011" -->
<!-- data key contains NA's, which are ignored -->
<!-- Over coverage. 6628 out of 6975 unmatched data records: S01006789, S01006790, S01006791, S01006792, S01006793, ... -->
<!-- Setting up the model -->
<!-- Collecting 20000 samples -->
<!--   |====================================================================================| 100% -->
<!-- Summarising results -->
<!--  finished in  133.3 seconds -->
<!-- > round(apply(indicators_aberdeen, 2, quantile, c(0.5, 0.025, 0.975)),3) -->
<!--       RCI2004 RCI2012 D2004 D2012 -->
<!-- 50%     0.218   0.187 0.385 0.360 -->
<!-- 2.5%    0.213   0.181 0.379 0.355 -->
<!-- 97.5%   0.224   0.193 0.390 0.366 -->
<!-- > ## Differences in RCI and D in 2011 - 2001 -->
<!-- > round(quantile(indicators_aberdeen[ ,2] - indicators_aberdeen[ ,1], c(0.5, 0.025, 0.975)),3) -->
<!--    50%   2.5%  97.5%  -->
<!-- -0.031 -0.040 -0.022  -->
<!-- > round(quantile(indicators_aberdeen[ ,4] - indicators_aberdeen[ ,3], c(0.5, 0.025, 0.975)),3) -->
<!--    50%   2.5%  97.5%  -->
<!-- -0.024 -0.032 -0.017  -->

@@

The following plot shows scatterplots of the ID proportions in each Scottish datazone in 2004, 2009 and 2012. Within these plots the blue line indicates parity between values on the x and y axis, and the red dashed line shows a nonlinear smoother function between the two ID proportions being plotted, and hence the level of deviation between the actual relationship over the two years and parity.

```{r show_temporal_dep, echo = FALSE, warning = FALSE, EVAL = TRUE}

@@

The following plot shows the change in the total population size as a function of distance from city centre. (The smoother line at distances less than 1km should likely be ignored for most cities as it involves extrapolation based on very few observations).

```{r plot_changetotalpop, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
totalpopchange_plot_for_city <- function(this_city){
  simd_2011_reweighted  %>% 
    ungroup() %>%  
    left_join(centre_distance, by = c("dz_2011" = "datazone")) %>% 
    filter(nearest_centre == this_city) %>%
    distinct() %>% 
    select(dz_2011, distance_to_centre, year, pop_total) %>% 
    mutate(distance_to_centre = distance_to_centre / 1000) %>% 
    spread(year, pop_total) %>% 
    mutate(change = (`2012` / `2004`) - 1) %>% 
    ggplot(., aes(x = distance_to_centre, y = change)) + 
    geom_point( alpha = 0.01) + 
    scale_x_log10(limits = c(0.5, 50), breaks = c(0.5, 1, 2, 5, 10, 20, 50)) + 
    scale_y_continuous(limits = c(-.5, .5)) +    
    stat_smooth() + 
    geom_vline(aes(xintercept = 1e+0)) + 
    geom_vline(aes(xintercept = 1.2e+1)) +
    geom_hline(aes(yintercept = 0), linetype = "dashed") + 
    labs(x = "Distance to centre (km)", y = "Change in total pop",
         title = this_city)
}

plot_grid(
  totalpopchange_plot_for_city("Aberdeen"),
  totalpopchange_plot_for_city("Dundee"),
  totalpopchange_plot_for_city("Edinburgh"),
  totalpopchange_plot_for_city("Glasgow"),
  nrow = 2
)
```

This plot suggests that there have been substantial increases in population sizes in Aberdeen, Edimburgh and Glasgow over this period, in datazones closer to the city centre, with the greatest changes occurring near the centre of Glasgow. In Aberdeen, and to a lesser extent for Dundee, there have also been increases in population shares living at distances of 12km and more from the city, suggesting greatest consistency in population sizes at intermediate distances.

@@

Alternative methods of reweighting SIMD values released at 2001 datazone levels could be explored, including those which make use of more sophisticated GIS-based methods to apportion values to 2011 datazones, and of approaches which use SIMD component values available at a higher level of spatial disaggregation (e.g. output areas) to recalculate and reapportion SIMD scores to 2011 datazone geographies.
